\textbf{Evaluating RISC-V Enclaves:}\\ Security Without the Slowdown?

### 2025-09-01 chapters/discussion.tex
last minute changes

To address overhead and scalability challenges, \emph{enclave multiplexing} offers a promising software-level strategy by consolidating multiple lightweight workloads into a single enclave. This reduces the number of PMP entries required by reusing them across tasks, optimizing resource usage and improving scalability. An example of a multiplexed enclave task dispatcher is shown in Listing~\ref{lst:enclave-mux}.
\vspace{8em}
\begin{lstlisting}[language=Python, caption={Multiplexed enclave task dispatcher.}, label={lst:enclave-mux}]
def enclave_dispatcher(task_id, input_data):
    if task_id == 1:
        return process_crypto(input_data)
    elif task_id == 2:
        return handle_key_store(input_data)
    elif task_id == 3:
        return run_diagnostics(input_data)
    else:
        raise ValueError("Unknown task")
\end{lstlisting}

To address overhead and scalability challenges, \emph{enclave multiplexing} offers a promising software-level strategy by consolidating multiple lightweight workloads into a single enclave. This reduces the number of PMP entries required by reusing them across tasks, optimizing resource usage and improving scalability. An example of a multiplexed enclave task dispatcher is shown in Listing~\ref{lst:enclave-mux}.

@@

In terms of software strategies, significant improvements can be made by enhancing the runtime management of enclaves, as outlined in Section~\ref{sec:synthesis} (Chapter~\ref{chap:discussion}). For example, reducing the frequency of context switches between the enclave and host operating system, and batching enclave invocations, could help mitigate the performance degradation caused by these transitions. Additionally, improving the enclave lifecycle management to dynamically allocate resources based on the current workload could lead to more efficient use of limited resources, further enhancing scalability.

Moreover, careful attention should be given to the design of memory allocation and management strategies in multi-tenant environments, where multiple enclaves may need to coexist on the same system (Section~\ref{sec:implications}, Chapter~\ref{chap:benchmarking}). In such scenarios, strategies such as enclave multiplexing—where multiple tasks or applications are combined within a single enclave—could significantly reduce the overall resource requirements. This approach, while potentially reducing the level of isolation between tasks, could help manage the limited PMP entries more effectively, enabling a larger number of enclaves to coexist within the same system.

>>>>

To address overhead and scalability challenges, \emph{enclave multiplexing} \cite{Occlum2020} offers a promising software-level strategy by consolidating multiple lightweight workloads into a single enclave. This reduces the number of PMP entries required by reusing them across tasks, optimizing resource usage and improving scalability. An example of a multiplexed enclave task dispatcher is shown in Listing~\ref{lst:enclave-mux}.
\vspace{8em}
\begin{lstlisting}[language=Python, caption={Multiplexed enclave task dispatcher.}, label={lst:enclave-mux}]
def enclave_dispatcher(task_id, input_data):
    if task_id == 1:
        return process_crypto(input_data)
    elif task_id == 2:
        return handle_key_store(input_data)
    elif task_id == 3:
        return run_diagnostics(input_data)
    else:
        raise ValueError("Unknown task")
\end{lstlisting}

To address overhead and scalability challenges, \emph{enclave multiplexing} \cite{Occlum2020} offers a promising software-level strategy by consolidating multiple lightweight workloads into a single enclave. This reduces the number of PMP entries required by reusing them across tasks, optimizing resource usage and improving scalability. An example of a multiplexed enclave task dispatcher is shown in Listing~\ref{lst:enclave-mux}.

@@

In terms of software strategies, significant improvements can be made by enhancing the runtime management of enclaves, as outlined in Section~\ref{sec:synthesis}. For example, reducing the frequency of context switches between the enclave and host operating system, and batching enclave invocations, could help mitigate the performance degradation caused by these transitions. Additionally, improving the enclave lifecycle management to dynamically allocate resources based on the current workload could lead to more efficient use of limited resources, further enhancing scalability.

Moreover, careful attention should be given to the design of memory allocation and management strategies in multi-tenant environments, where multiple enclaves may need to coexist on the same system (Section~\ref{sec:implications}). In such scenarios, strategies such as enclave multiplexing—where multiple tasks or applications are combined within a single enclave—could significantly reduce the overall resource requirements. This approach, while potentially reducing the level of isolation between tasks, could help manage the limited PMP entries more effectively, enabling a larger number of enclaves to coexist within the same system.

### 2025-09-01 preamble.tex
last minute changes

\UzLThesisSetup{
Masterarbeit,
Logo-Dateiname        = {uzl-iti-logo-uzl.pdf},
Verfasst              = {am}{Institut für Technische Informatik},
Titel auf Deutsch     = {
    Bewertung von RISC-V-Enklaven: Leistungsbenchmarks und bewährte Konfigurationspraktiken
}, 
Titel auf Englisch    = {
    Evaluating RISC-V Enclaves: Performance Benchmarks and Configuration Best Practices 
},
Autor                 = {Basil Ugbomoiko},
Betreuerin            = {Dr.-Ing. Saleh Mulhem},
Mit Unterstützung von = {Henrik Strunck, M.Sc},
Studiengang           = {IT Security},
Datum                 = {01. September 2025},
Abstract = {
  This thesis evaluates the performance of Keystone enclaves, an open-source Trusted Execution Environment (TEE) for RISC-V platforms. Keystone isolates sensitive workloads to enable secure computation, and its performance is assessed across CPU cores, cache sizes, memory allocations, and enclave configurations. Using Dhrystone, CoreMark, and cryptographic workloads such as Kyber—a post-quantum key encapsulation mechanism—this work measures execution time, context switch overhead, memory latency, and CPU utilization. The results reveal critical bottlenecks and provide best-practice configuration guidelines for compute-bound, memory-intensive, mixed, and post-quantum workloads, supporting efficient and secure deployment of Keystone enclaves on RISC-V architectures.
},
Zusammenfassung = {
  Diese Arbeit bewertet die Leistung von Keystone-Enklaven, einer Open-Source Trusted Execution Environment (TEE) für RISC-V-Plattformen. Keystone isoliert sensible Workloads zur sicheren Berechnung, wobei Leistungseinflüsse durch CPU-Kerne, Cache-Größe, Speicherzuweisung und Enklavenkonfiguration untersucht werden. Mithilfe von Dhrystone, CoreMark und kryptografischen Workloads wie Kyber—einem postquantischen Schlüssel-Kapselungsverfahren—werden Ausführungszeit, Kontextwechsel-Overhead, Speicherlatenz und CPU-Auslastung gemessen. Die Ergebnisse zeigen zentrale Engpässe auf und liefern Konfigurationsrichtlinien für rechenintensive, speicherintensive, gemischte und postquantische Workloads, um Keystone-Enklaven effizient und sicher auf RISC-V-Architekturen einzusetzen.
},
Acknowledgements = {
    I would like to sincerely thank Daniel K. for his insightful guidance and clarifications, which added immense value to this work.
    I am also grateful to Tom K. for providing the initial log analysis script and for his valuable input on effective visualization approaches.
    My heartfelt thanks go to my parents for their patience, encouragement, and the opportunities they have provided.  
    I deeply appreciate my supervisors, Dr.-Ing. Saleh M. and Henrik S. (M.Sc), for their guidance, constructive feedback, and steadfast support throughout this project.  
    Finally, I extend my gratitude to my fellow Lübeckers and all others who contributed in various ways; their support, though too numerous to mention individually, was invaluable.
},
Numerische Bibliographie
}

>>>>

\UzLThesisSetup{
Masterarbeit,
Logo-Dateiname        = {uzl-iti-logo-uzl.pdf},
Verfasst              = {am}{Institut für Technische Informatik},
Titel auf Deutsch     = {
    Bewertung von RISC-V-Enklaven: Leistungsbenchmarks und bewährte Konfigurationspraktiken
}, 
Titel auf Englisch    = {
    Evaluating RISC-V Enclaves: Performance Benchmarks and Configuration Best Practices 
},
Autor                 = {Basil Ugbomoiko},
Betreuerin            = {Dr.-Ing. Saleh Mulhem},
Mit Unterstützung von = {Henrik Strunck, M.Sc.},
Studiengang           = {IT Security},
Datum                 = {31. August 2025},
Abstract = {
  This thesis evaluates the performance of Keystone enclaves, an open-source Trusted Execution Environment (TEE) for RISC-V platforms. Keystone isolates sensitive workloads to enable secure computation, and its performance is assessed across CPU cores, cache sizes, memory allocations, and enclave configurations. Using Dhrystone, CoreMark, and cryptographic workloads such as Kyber—a post-quantum key encapsulation mechanism—this work measures execution time, context switch overhead, memory latency, and CPU utilization. The results reveal critical bottlenecks and provide best-practice configuration guidelines for compute-bound, memory-intensive, mixed, and post-quantum workloads, supporting efficient and secure deployment of Keystone enclaves on RISC-V architectures.
},
Zusammenfassung = {
  Diese Arbeit bewertet die Leistung von Keystone-Enklaven, einer Open-Source Trusted Execution Environment (TEE) für RISC-V-Plattformen. Keystone isoliert sensible Workloads zur sicheren Berechnung, wobei Leistungseinflüsse durch CPU-Kerne, Cache-Größe, Speicherzuweisung und Enklavenkonfiguration untersucht werden. Mithilfe von Dhrystone, CoreMark und kryptografischen Workloads wie Kyber—einem postquantischen Schlüssel-Kapselungsverfahren—werden Ausführungszeit, Kontextwechsel-Overhead, Speicherlatenz und CPU-Auslastung gemessen. Die Ergebnisse zeigen zentrale Engpässe auf und liefern Konfigurationsrichtlinien für rechenintensive, speicherintensive, gemischte und postquantische Workloads, um Keystone-Enklaven effizient und sicher auf RISC-V-Architekturen einzusetzen.
},
Acknowledgements = {
    I would like to sincerely thank Daniel K. for his insightful guidance and clarifications, which added immense value to this work.
    I am also grateful to Tom K. for providing the initial log analysis script and for his valuable input on effective visualization approaches.
    My heartfelt thanks go to my parents for their patience, encouragement, and the opportunities they have provided.  
    I deeply appreciate my supervisors, Dr.-Ing. Saleh M. and Henrik S. (M.Sc.), for their guidance, constructive feedback, and steadfast support throughout this project.  
    Finally, I extend my gratitude to my fellow Lübeckers and all others who contributed in various ways; their support, though too numerous to mention individually, was invaluable.
},
Numerische Bibliographie
}
